### Deciding the quality of a work out.


This is the machine learning algorithm that I generated to determine if a workout was of high quality. 


#### First we load the libraries.
```{r create directory, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}
library(tibble)
library(tidyverse)
library(lubridate)
library(dplyr)
library(readr)
library(caret)
if(!file.exists("data")){
  dir.create("data")
}
```

#### Next we will load the data and store it under a directory './data'
```{r get data, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}
fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileUrl,destfile="./data/pmltraining.csv",method="curl")
fileUrl2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl2,destfile="./data/pmltesting.csv",method="curl")
list.files("./data")
dateDownloaded <- date()
dateDownloaded
```


#### Next we will load the data into a dataframe.
``` {r read data, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}
measurements <- (read_delim(file = "./data/pmltraining.csv", col_names = TRUE,
                      delim = ","))
```

#### Next the data needs to be cleaned.


In the initial data set there are 160 variables in each observation.

For my machine learning, i only looked at the gyros_, accel_, and magnet_ fields for the dumbell, forearm, arm, and wrist fields.  In addition I also used the timestamp and num_window field as well.

This reduced my observations from 160 to 38.


``` {r cleaning the data, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}

measurements <- bind_cols(measurements[3:7],measurements[37:45], measurements[60:68], measurements[113:120], measurements[151:160])
measurements <- add_column(measurements, cvt_time = as_datetime(measurements$raw_timestamp_part_1 + measurements$raw_timestamp_part_2))
measurements <- bind_cols(measurements[6:42], measurements[5])
head(measurements)
```
#### Create Training and Test Sets


Next I created a training and a test set.  The training set is about 60% of the data, and the test set is about 40% of the data.


``` {r create training and test sets, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}

inTrain <- createDataPartition(y=measurements$classe,
                              p=.6, list=FALSE)
training <- measurements[inTrain,]
test <-   measurements[-inTrain,]
```

#### Exploratory Data Analysis.


Below are several plots that show there is some relationship between the accel_, gyros_, and magnet_ values and any one of the 4 areas where the sensors were worn.  I will use this as the basis for my machine learning algorithm.  

I also got concurance on this opinion from the paper that was referenced below:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: <a href="http:/groupware.les.inf.puc-rio.br/har#weight_lifting_exercises#ixzz4TkQaPnXd">Source</a>


``` {r explorator data analysis, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}

featurePlot(x=training[,c("accel_belt_x", "accel_belt_y", "accel_belt_z")],
             y = factor(training$classe),
            plot="pairs", no.legend=F)

featurePlot(x=training[,c("gyros_belt_x", "gyros_belt_y", "gyros_belt_z")],
            y = factor(training$classe),
            plot="pairs", no.legend=F)

featurePlot(x=training[,c("magnet_belt_x", "magnet_belt_y", "magnet_belt_z")],
            y = factor(training$classe),
            plot="pairs", no.legend=F)

```
#### Model Generation


Next I will train my model using the training data set.  For this I used a random forest, initially I did try this with a boosting algorithm but I later found that the random forest algorithm provided a higher level of accuracy against the training set. 


``` {r create model1, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}

modelFit <- train(classe ~. -num_window,data=training, method="rf",  allowParrallel = TRUE)

modelFit
```


As you can see here is that the accuracy of this model is about 96%.


#### Future Predictions


Below are my future predictions using the separate training set that I generated at the beginning.


``` {r predicting new values, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE}
pred <- predict(modelFit,test);

test$predRight <- pred==test$classe
table(pred,test$classe)

qplot(cvt_time,num_window,colour=predRight,data=test,main="newdata Predictions")
```

#### Conclusion


As you can see from the above table, my algorithm was highly effective at identifying the classe of workout given the gyros_, accel_, and magnet_ observations for all for sources (dumbbell, forearm, arm, belt)